Le titre

### 2022-10-16 Manuscrit Thèse/chapter 4/Chapter 4.tex
tata

\DeclareUnicodeCharacter{0308}{HERE!HERE!} % je vais pas m'énerver, mais un peu quand meme
\title{}
\begin{document}
\chapter{NV-NV CR under transverse or low fields: application to magnetometry}

In the last chapter we saw how dense ensemble of NV centers have an intrinsic depolarization mechanism mediated by the NV-NV dipole interaction. In this chapter we will see how this feature can be exploited to perform magnetometry.

Using NV-NV CR for magnetometry in non-zero magnetic field has already been proposed \citep{akhmedzhanov2017microwave, akhmedzhanov2019magnetometry}. We will focus here on the low to zero magnetic field region, which presents practical advantages as the relative orientation of the diamond with the magnetic field does not need to be controlled, but theoretical complications due to the more complex physics of the NV centers under low magnetic field.

This chapter is constructed as follows: We will first describe the physics of NV centers under low or transverse magnetic field (the physics is similar in both cases), we will then show both theoretically and experimentally that the NV-NV CR is modified for low and transverse magnetic field. We will identify different mechanisms by which the relaxation is modified and we will try to quantify their relative importance.

We will then focus on the low field depolarization magnetometry (LFDM) protocol which exploits the previously studied NV-NV CR at low magnetic field. We will start by giving a short overview of NV ensemble magnetometry before characterizing the LFDM protocol and comparing it to other magnetometry protocols. Finally we will give some perspectives on how to improve LFDM and on potential applications.

@@

\section{NV spin Hamiltonian under low and transverse fields}

Before looking at the NV-NV CR in the low or transverse field regime, we first need to consider how the general NV physics is modified under those regimes, and in particular we need to look at the modifications of the spin Hamiltonian and the change in the eigenstates compared to the high magnetic field regime.

\subsection{NV spin Hamiltonian in zero external magnetic field}
\label{sec. Hamltonian 0 B}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{Figures/ESR_simus}
\caption{Simulations of inhomogeneous zero field ODMR when sampling various parameters. a) Simulation when sampling each components of the magnetic field over a Gaussian of deviation $\sigma=2\ \rm G$. b) Simulation when sampling each components of the strain tensor $\bar{\bar{\varepsilon}}$ over a Gaussian of deviation $\sigma=2\cdot 10^{-4}$. c) Simulation when sampling each components of the electric field over a Gaussian of deviation $\sigma=2\cdot 10^{5}\ \rm{V/cm}$. d) Experimental ODMR spectrum in zero external field taken on sample ADM-150-2}
\label{simus ESR}
\end{figure}

In the absence of external magnetic field, we have to take into account other elements which would otherwise be of second order in the spin Hamiltonian. These elements are: the random local magnetic fields caused by paramagnetic impurities, the local electric field caused by charged impurities, and the crystal strain \citep{doherty2012theory, udvarhelyi2018spin, mittiga2018imaging}. The hyperfine splitting due to nearby nuclei will be considered separately, although to a large extent it behaves like a local magnetic field.

Due to the large zero field splitting $D=2870\ \rm MHz$ between the $\ket{0}$ and $\ket{\pm 1}$ states, we will consider the $\ket{0}$ state to always be an eigenstate of the spin Hamiltonian under zero external field (which is equivalent to say that we neglect the terms in $\ket{0}\bra{\pm 1}$ in the spin Hamiltonian). The problem is then reduced to the $\{ \ket{-1}, \ket{+1} \}$ subsystem.

@@

Importantly, as pointed in \citep{mittiga2018imaging} we notice that both the electric field and the strain have a \textit{shifting} component ($f_\parallel$ and $g_\parallel$) which shifts equally both eigenstates of the Hamiltonian, and a \textit{splitting} component ($f_\perp$ and $g_\perp$) which splits in energy the two eigenstates.

The main difference between the electric field and the strain is in the numerical prefactors of these components: for the electric field, the splitting parameter $d_\perp$ is $\sim 50$ times higher than the shifting parameter $d_\parallel$, which will result on average to a strong energy split without much shifting. For the strain however, the splitting parameters $h_{15}$ and $h_{16}$ are only $\sim 3$ times higher than the shifting parameters $h_{43}$ and $h_{41}$. The shift in energy will therefore tend to blur the energy split when averaging over a large number of spins.

Fig. \ref{simus ESR} shows a simulation of how each parameters of the spin Hamiltonian - local magnetic field, local electric field and strain - affects the zero external field ODMR profile. To do these simulations, I sampled each parameters separately $10^6$ times and plotted the histogram of the two eigenvalues of the Hamiltonian written in eq. \ref{Hamiltonien pm1}. Fig. \ref{simus ESR}-d) shows an experimental zero field ODMR spectrum, typical of what we observe with dense NV ensembles.

Experimentally, almost all our samples show the characteristic two bumps in zero external field ODMR. Given the simulation results, the only parameter that can give rise to this shape is the local electric field.

We should note that another potential explanation for the splitting of the $\ket{\pm 1}$ states could be the presence of a residual magnetic field, not randomly distributed among all the spins but homogeneous on the whole sample. In particular, we do not shield for the earth magnetic field in our experiment. The splitting caused by the earth magnetic field however has a value $\gamma B_{\rm earth} \approx 1\ \rm MHz$, which is $\sim 4$ times weaker than the splitting we typically observe.

We will therefore consider that the low external magnetic field spin Hamiltonian of our samples is dominated by the local electric field, and more specifically by the transverse electric field $E_\perp\equiv E_x + i E_y$ given the ratio between $d_\perp$ and $d_\parallel$.

We will then adopt the following simplified Hamiltonian for the zero external field regime:
\begin{equation}
\mathcal{H}=\begin{pmatrix}
D&0&d_\perp E_\perp^* \\
0&0&0 \\
d_\perp E_\perp &0&D
\end{pmatrix},
\end{equation}
whose eigenvectors are $\ket{0}$ and$\ket{\pm}$ of eigenvalues 0 and $D\pm d_\perp \abs{E_\perp}$, where $\ket{\pm}$ are defined as:
\begin{equation}
\ket{\pm}=\frac{\ket{+1}\pm e^{-i\phi_E}\ket{-1}}{\sqrt{2}},
\end{equation}
where $\tan(\phi_E)=E_y/E_x$.

@@

We now turn to the case of purely transverse magnetic field with respect to the NV axis, i.e. $\mathbf{B}=B_x \hat{e}_x + B_y \hat e_y$, and more specifically to the regime where $d_\perp E_\perp < \frac{(\gamma_e B_\perp)^2}{D} << D$. In practice, this generally corresponds to $20\ \rm G \lesssim B_\perp \lesssim 200\ \rm G$.

@@

In conclusion, whenever the spin Hamiltonian is dominated by a transverse field, either electric or magnetic, we can consider that the eigenstates of the spin Hamiltonian to be $\ket{0}, \ket{-}$ and $\ket{+}$, whereas when the Hamiltonian is dominated by the longitudinal magnetic field, its eigenstates are $\ket{0}, \ket{-1}$ and $\ket{+1}$.

@@

A consequence of the change in the Hamiltonian eigenstates from the $\{ \ket{0}, \ket{\pm 1} \}$ to the $\{ \ket{0}, \ket{\pm} \}$ basis is that the Hamiltonian eigenvalues are sensitive to different part of the environment. The energy of the $\ket{\pm 1}$ states depend linearly on the longitudinal magnetic field, and to the second  order on the transverse electric and magnetic fields. On the contrary, the $\ket{\pm}$ states depend linearly on the transverse electric field, and to the second order on the longitudinal and transverse magnetic field \footnote{The transverse magnetic field is always of second order for weak magnetic field because $D\gg \gamma B_\perp$.}.

The shape of the ODMR lines (i.e. the spectral response of the spins) is therefore dominated by the magnetic noise and the hyper-fine coupling in the $\{ \ket{0}, \ket{\pm 1} \}$ basis, and by the electric noise in the $\{ \ket{0}, \ket{\pm} \}$ basis. Further details on the low magnetic field physics of the NV center are available in \citep{jamonneau2016competition} and on the purely transverse magnetic field regime in \citep{qiu2021nuclear, qiu2022nanoscale}.

@@

\section{NV-NV CR in the low magnetic field regime}
In this part, we will study the cross-relaxation between NV centers under low magnetic field, whose spin Hamiltonian is dominated by a transverse electric field. Understanding the various mechanisms operating in this regime will be important for the magnetometry protocol presented after.

\subsection{Experimental observations}
\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{Figures/scan_1x1x1x1}
\caption{Low field depolarization on sample ADM-150-1 for $\mathbf{B}$ in an arbitrary direction. a) ODMR spectrum for B=0. b) ODMR spectrum for B=60 G. c) PL contrast as a function of the external magnetic field. d) $1/T_1^{\rm dd}$ as a function of the external magnetic field. A $T_1$ measurement was recorded for each magnetic field value and fitted accorded to the protocol described in [REF] with $T_1^{\rm ph}=5\ \rm ms$}
\label{scan 1x1x1x1}
\end{figure}
We will start by showing that NV-NV CR behaves differently in the low magnetic field regime compared to the longitudinal field dominated regime which we studied in the last chapter.

The main issue with studying NV-NV CR in low to zero magnetic field is that there are many competing effects happening simultaneously, with few buttons to adjust in order to isolate each effects.

Fig. \ref{scan 1x1x1x1}-c) and d) show the evolution of the NV PL and stretched lifetime $T_1^{\rm dd}$, defined in the last chapter [REF], as the magnetic field is scanned from 0 to 60 G. The ODMR at the initial and final magnetic fields are shown in Fig. \ref{scan 1x1x1x1}-a) and b).

While it is clear that the spin lifetime as well as the PL increases with the magnetic field, there is no clear indication that this is because of the specificity of the low field region. Indeed, the most likely explanation in this case is that the four classes of NV centers get split apart as the magnetic field increases, which reduces the density of resonant fluctuators for each NV centers.

We can note however that the PL is not an exact mirror of the spin dynamics: while the low field drop in PL is indeed associated with a reduced spin lifetime, the drop in PL at higher magnetic field comes from the states mixing induced by the transverse magnetic field, and is not associated with a modification of the spin dynamics (to the first order).

@@

Fig. \ref{scan 100} presents a way to circumvent this issue: by applying the magnetic field along the [100] crystalline axis, we can make sure that the four classes of NV centers always stay resonant regardless of the magnetic field amplitude.

We can notice that there still is a decrease of both the PL and $T_1^{\rm dd}$ in low field, although considerably smaller than the previous case: in Fig. \ref{scan 1x1x1x1}, $T_1^{\rm dd}$ was reduced by a factor of $\sim$ 8 in zero field, whereas in Fig. \ref{scan 100}, it was only reduced by a factor of $\sim$ 1.2. The main reason for the PL and $T_1$ drop in Fig. \ref{scan 1x1x1x1} was indeed the co-resonance between the four classes.

Nevertheless, the fact that there is a drop in zero field when $\mathbf{B} \parallel [100]$ cannot be explained if we consider only the inter-class resonances. This means that there are some additional depolarization mechanisms which are proper to the zero field region. We should also note that, while the zero-field PL contrast is bigger in Fig. \ref{scan 1x1x1x1}-c) than in Fig. \ref{scan 100}-c), the PL slope, which is the limiting factor for sensing ability, is actually similar in both cases.

We can also note a drop in PL and a corresponding increase of $1/T_1^{\rm dd}$ for $B \sim 20\ \rm G$ which corresponds to the NV$-^{13}$C-NV cross-relaxation discussed in [REF]

\subsection{Theory of the low field depolarization}
\label{sec causes zero field}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{Figures/Shema_summary_theory}
\caption{Visual summary of the different depolarization effects in low magnetic field and their predicted depolarization rate}
\label{summary_theory}
\end{figure}

>>>>

\DeclareUnicodeCharacter{0308}{HERE!HERE!} % je vais pas m'énerver, mais un peu quand meme
\title{}
\begin{document}
\chapter{NV-NV cross-relaxation at low fields: application to magnetometry}

In the last chapter we have identified a dipole-dipole relaxation mechanism within dense NV centers ensemble, and we have studied its influence in non-zero magnetic field. In this chapter we will cover the low to zero magnetic field region and see how the depolarization at low magnetic field can be used for magnetometry.

This chapter is constructed as follows: We will first describe in sec. \ref{sec 4.1} the physics of NV centers under low or transverse magnetic field. In sec. \ref{sec causes zero field}
we will present the different zero field depolarization causes that we have identified as well as their theoretical impact on the NV relaxation rates according to the fluctuator model. In sec. \ref{sec 4.3} we will confront the predictions of the fluctuator model with experiments under low and transverse magnetic field.

We will then focus on the low field depolarization magnetometry (LFDM) protocol which exploits the previously studied NV-NV CR at low magnetic field. We will start by giving a short overview of NV ensemble magnetometry in sec. \ref{sec 4.4} before characterizing the LFDM protocol in sec. \ref{sec LFDM} and comparing it to other magnetometry protocols in sec. \ref{sec 4.6}. Finally we will give some perspectives on how to improve LFDM and on potential applications.

@@

\section{NV spin Hamiltonian under low and transverse fields}
\label{sec 4.1}
We will start this chapter by studying the single NV physics under low and transverse magnetic field. On of the main difference with the previously studied regime is that the eigenstates of the Hamiltonian are no longer $\{ \ket{0}, \ket{+1}, \ket{-1} \}$, which will then in turn modify the dipole-dipole interaction between the NV centers.

We will first describe the spin Hamiltonian of the NV center under low magnetic field and determine the dominant factors in our samples, then do the same for transverse magnetic field, and finally investigate the modification of the ODMR lineshapes in those regimes.

\subsection{NV spin Hamiltonian in zero external magnetic field}
\label{sec. Hamltonian 0 B}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{Figures/ESR_simus}
\caption{Monte-Carlo simulations of inhomogeneous zero field ODMR for different noise sources. a) Magnetic noise: each of the $\mathbf{B}$ components is sampled over a Gaussian of deviation $\sigma=2\ \rm G$. b) Strain noise: each of the components of the strain tensor $\bar{\bar{\varepsilon}}$ is sampled over a Gaussian of deviation $\sigma=2\cdot 10^{-4}$. c) Electric noise: each of the $\mathbf{E}$ components is sampled over a Gaussian of deviation $\sigma=2\cdot 10^{5}\ \rm{V/cm}$. d) Experimental ODMR spectrum in zero external field taken on sample ADM-150-2}
\label{simus ESR}
\end{figure}

In the absence of a strong external magnetic field, the dominant terms of the NV spin Hamiltonian originate from local noises. These noises are: the random local magnetic fields caused by paramagnetic impurities, the local electric field caused by charged impurities, and the crystal strain \citep{doherty2012theory, udvarhelyi2018spin, mittiga2018imaging}.

Due to the large zero field splitting $D=2870\ \rm MHz$ between the $\ket{0}$ and $\ket{\pm 1}$ states, we will consider the $\ket{0}$ state to always be an eigenstate of the spin Hamiltonian under zero external field. The problem is then reduced to the $\{ \ket{-1}, \ket{+1} \}$ subsystem.

@@

%Importantly, as pointed in \citep{mittiga2018imaging} we notice that both the electric field and the strain have a \textit{shifting} component ($f_\parallel$ and $g_\parallel$) which shifts equally both eigenstates of the Hamiltonian, and a \textit{splitting} component ($f_\perp$ and $g_\perp$) which splits in energy the two eigenstates. 
%
%The main difference between the electric field and the strain is in the numerical prefactors of these components: for the electric field, the splitting parameter $d_\perp$ is $\sim 50$ times higher than the shifting parameter $d_\parallel$, which will result on average to a strong energy split without much shifting. For the strain however, the splitting parameters $h_{15}$ and $h_{16}$ are only $\sim 3$ times higher than the shifting parameters $h_{43}$ and $h_{41}$. The shift in energy will therefore tend to blur the energy split when averaging over a large number of spins.

Fig. \ref{simus ESR}-a) b) and c) shows the result of Monte-Carlo simulations of the ODMR of an ensemble of NV centers for different noise sources. The simulation work as follow: one of the three possible noise source (magnetic field, electric field or strain) is sampled randomly while the two other are kept at zero. The eigenvalues of the spin Hamiltonian given in eq. \ref{Hamiltonien pm1} are then computed and the process is repeated $10^6$ times to create the histograms shown in Fig. \ref{simus ESR}.

Fig. \ref{simus ESR}-d) shows an experimental zero-field ODMR which is typical of the samples used in this manuscript. In particular we can notice a central splitting of the spectrum. This splitting is only present in the simulation when the dominant noise is the electric field. We then conclude that the dominant noise source in our sample (in the absence of external magnetic field) is the local electric field.

The authors of \citep{mittiga2018imaging} came to the same conclusion for the samples with dense NV ensemble that they studied by using a more sophisticated simulation model.

%We should note that another potential explanation for the splitting of the $\ket{\pm 1}$ states could be the presence of a residual magnetic field, not randomly distributed among all the spins but homogeneous on the whole sample. In particular, we do not shield for the earth magnetic field in our experiment. The splitting caused by the earth magnetic field however has a value $\gamma B_{\rm earth} \approx 1\ \rm MHz$, which is $\sim 4$ times weaker than the splitting we typically observe.

Given that $d_\perp \approx 50 d_\parallel$, we can consider that the dominant electric field noise comes from the transverse electric field. We will then adopt the following simplified Hamiltonian for the zero external magnetic field regime:
\begin{equation}
\mathcal{H}=\begin{pmatrix}
D&0&d_\perp E_\perp^* \\
0&0&0 \\
d_\perp E_\perp &0&D
\end{pmatrix},
\end{equation}
whose eigenvectors are $\ket{0}$ and$\ket{\pm}$ of eigenvalues 0 and $D\pm d_\perp \abs{E_\perp}$, where $\ket{\pm}$ are defined as:
\begin{equation}
\ket{\pm}=\frac{\ket{+1}\pm e^{-i\phi_E}\ket{-1}}{\sqrt{2}},
\end{equation}
where $\tan(\phi_E)=E_y/E_x$.

@@

We now turn to the case of a magnetic field that is orthogonal to the NV axis, i.e. $\mathbf{B}=B_x \hat{e}_x + B_y \hat e_y$, and more specifically to the regime where $d_\perp E_\perp < \frac{(\gamma_e B_\perp)^2}{D} << D$. In practice, this generally corresponds to $20\ \rm G \lesssim B_\perp \lesssim 200\ \rm G$.

@@

In conclusion, whenever the spin Hamiltonian is dominated by the local electric field or by the transverse magnetic field, we can consider that the eigenstates of the spin Hamiltonian to be $\ket{0}, \ket{-}$ and $\ket{+}$, whereas when the Hamiltonian is dominated by the longitudinal magnetic field, its eigenstates are $\ket{0}, \ket{-1}$ and $\ket{+1}$.

@@

A consequence of the change in the Hamiltonian eigenstates from the $\{ \ket{0}, \ket{\pm 1} \}$ to the $\{ \ket{0}, \ket{\pm} \}$ basis is that the Hamiltonian eigenvalues are sensitive to different part of the environment. The energy of the $\ket{\pm 1}$ states depends linearly on the longitudinal magnetic field, and quadratically on the transverse electric and magnetic fields. On the contrary, the $\ket{\pm}$ states depend linearly on the transverse electric field, and quadratically on the longitudinal and transverse magnetic field \footnote{The transverse magnetic field is always of second order for weak magnetic field because $D\gg \gamma B_\perp$.}.

The shape of the ODMR lines (i.e. the spectral response of the spins) is therefore dominated by the magnetic noise and the hyper-fine coupling in the $\{ \ket{0}, \ket{\pm 1} \}$ basis, and by the electric noise in the $\{ \ket{0}, \ket{\pm} \}$ basis, as we established in the previous part. Further details on the low magnetic field regime are available in \citep{jamonneau2016competition} and on the purely transverse magnetic field regime in \citep{qiu2021nuclear, qiu2022nanoscale}.

@@

\section{Theory of the low field depolarization}
\label{sec causes zero field}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{Figures/Shema_summary_theory_2}
\caption{Summary of the different depolarization effects in low magnetic field and their predicted depolarization rate. The ODMR spectra were taken on sample ADM-150-1}
\label{summary_theory}
\end{figure}

We will focus on this section on the mechanisms responsible for the depolarization of dense NV ensemble at low magnetic fields. We have identified four such mechanisms that lead to an increase in the resonant dipole-dipole coupling between NV and fluctuators, which in turn results in an increased spin depolarization rate.

Fig. \ref{summary_theory} summarizes these four mechanisms along with the theoretical increase in the depolarization rate caused by each of them. We will detail below the physical origin of these four mechanisms and the calculations that lead to the values presented in Fig. \ref{summary_theory}.

\subsection{Four classes degeneracy}

The first and main contribution to the low field depolarization is the degeneracy between the four classes as the magnetic field decreases. Following the arguments presented in the last chapter, the degeneracy between the four classes of NV centers causes in theory an increase in the relaxation rate by a factor $\sim 43$. This contribution is not specific to the zero field region, as we can observe the same resonance condition when $\mathbf{B} \parallel [100]$.

\subsection{Eigenstates modification}

A consequence of this change is the absence of diagonal terms  in the dipole-dipole Hamiltonian written in the $\{ \ket{0}, \ket{+},\ket{-} \} \times \{ \ket{0}, \ket{+},\ket{-} \}$ basis, which explains why the NV centers states are protected from magnetic noises at low magnetic field \citep{jamonneau2016competition}.

Another consequence is a slight modification of the flip-flop rates. Once averaged over all possible orientation and position, as detailed in appendix [REF], this modification of the flip-flop rates amount for an increase of $\Gamma_1$ by about $20\ \%$ in the $\{ \ket{0}, \ket{+},\ket{-} \}$ basis compared to the $\{ \ket{0}, \ket{+1},\ket{-1} \}$ basis.
%The first explanation is the modification of the dipole-dipole interaction caused by the change of the NV Hamiltonian eigenbasis from $\{ \ket{0}, \ket{\pm 1} \}$ when $\mathbf{B} \neq 0$ to $\{ \ket{0}, \ket{\pm} \}$ when $\mathbf{B} = 0$.
%
%This modification arise from the new form of the dipole-dipole Hamiltonian in the $\{ \ket{0}, \ket{+}, \ket{-} \} \times \{ \ket{0}, \ket{+}, \ket{-} \}$ basis. We justify this change of basis, where we only considered the single NV Hamiltonian instead of the full two-spins Hamiltonian, by the fact that we are in the weak coupling regime, where $\expval{\mathcal{H}_{dd}} \approx 50\ \rm kHz \ll \frac{1}{2\pi T_2^*} \approx 5\ \rm MHz$ meaning that we can treat the dipole-dipole interaction perturbatively. To compute the decay rate with these new eigenstates, we now need to consider the $\mel{0,\pm}{\mathcal{H}_{dd}}{\pm ,0}$ matrix elements instead of the $\mel{0,\pm 1}{\mathcal{H}_{dd}}{\pm 1,0}$ ones.
%
%The computation of the decay rates in the new eigenbasis involve the averaging of these matrix elements which is detailed in appendix [REF]. The computation in this case is complicated by the fact that the transverse field (either $\mathbf{E}$ or $\mathbf{B}$) responsible for the splitting of the $\ket{\pm}$ levels breaks the Hamiltonian symmetry in the $(xy)$ plane. This means that the dipole-dipole coupling between two spins will depend on their relative $x$ and $y$ axis, defined by the local transverse field, on top of the relative $z$ axis defined by the NV axis. 
%
%We therefore need to make an assumption on the distribution of the transverse field in the sample. In zero external magnetic field where the dominant transverse field comes from randomly spaced charged impurities, we can expect the $x$ and $y$ axes to be randomly sampled in their respective $(xy)$ plane. However, since the NV-fluctuator CR is dominated by the closest neighbor of each spin (due to the $1/r^6$ scaling in eq. [REF]), there could still be local correlations in the transverse field felt by the NV and its nearest fluctuator.
%
%We then computed the decay rates for the two extreme cases: first, we consider that the $x$ and $y$ axis of each spin is randomly sampled, which correspond to a correlation length of the transverse field $l_c=0$, and secondly we considered the case where the dominant transverse field as a fixed orientation in the whole sample corresponding to $l_c=\infty$.
%
%We found for $\mathbf{B}=0$ that the expected decay rate was $\Gamma_1=51.4 \Gamma_0^{\rm th}$ if $l_c=0$ and $\Gamma_1=55.0 \Gamma_0^{\rm th}$ if $l_c=\infty$. $\Gamma_0^{\rm th}$ has the same definition as in table [REF], it is the expected decay rate for an isolated class in the $\{ \ket{0}, \ket{\pm 1} \}$ basis.
%
%In both cases, this is a moderate ($\sim 20 \%$) increase compared to $\Gamma_1=42.8 \Gamma_0^{\rm th}$ which we previously found for $\mathbf{B} \parallel [100]$, where all four classes were degenerate but the Hamiltonian eignebasis was $\{ \ket{0}, \ket{\pm 1} \}$. The change in the Hamiltonian eigenbasis for low field is therefore a possible candidate to explain the low field depolarization in Fig. \ref{scan 100}

\subsection{Double flips}

The double flip processes have been previously discussed in sec. [REF] and naturally arise from the dipole-dipole Hamiltonian. The ones that we consider here couple the states $\ket{0, \pm 1}$ and $\ket{\mp 1, 0}$ (in contrast with the flip-flop which couple the states $\ket{0, \pm 1}$ and $\ket{\pm 1, 0}$). These terms are generally not resonant because of the Zeeman splitting between the $\ket{+1}$ and $\ket{-1}$ states.

In the low magnetic field regime however, we have to consider the new eigenstates $\ket{+}$ and $\ket{-}$ and the two-qubits states $\ket{0, \pm}$ and $\ket{\mp, 0}$. These terms are still not fully resonant because of the splitting between the $\ket{+}$ and $\ket{-}$ states caused by the local electric field, but they are close enough in energy that the double flip processes are not entirely quenched.

The computation of the depolarization rate including the double flips is detailed in appendix [REF]. To obtain a numerical value, we need to input the experimental values of the dipole-dipole interaction range \footnote{By interaction range we mean the value to be compared to the spectral detuning $\Delta \nu$ between the two groups. This is $2\gamma_f+\Gamma_{NV}+\Gamma_f$ in eq. [REF]} which we measured to be $\Gamma=8.8\ \rm MHz$ in Fig. [REF], as well as the average splitting $\Delta \nu$ between the $\ket{+}$ and $\ket{-}$ states. We typically measure $\Delta \nu \approx 8\ \rm MHz$ on type Ib HPHT samples.

For the value presented in Fig. \ref{simus ESR}, we inputted $\Gamma=\Delta \nu=8.8\ \rm MHz$.
%\begin{table}[htbp]
%\centering
%\caption{Simulated depolarization rate for flip-flops and double flips in zero magnetic field}
% \label{table double flips}
%\begin{tabular}{c|cc}
%%\toprule
%{} & $l_c=0$ & $l_c=\infty$ \\
%\midrule
%$\bar{R}^2=1$ & $126.8 \Gamma_0^{\rm th}$ & $126.8 \Gamma_0^{\rm th}$ \\
%$\bar{R}^2=0.5$ & $93.7 \Gamma_0^{\rm th}$ & $95.4 \Gamma_0^{\rm th}$ \\
%$\bar{R}^2=0$ & $51.4 \Gamma_0^{\rm th}$ & $55 \Gamma_0^{\rm th}$ \\
%%\bottomrule
%\end{tabular}
%\end{table}
%
%Another effect which happens only at low magnetic field is the possibility of double flip between the NV center and the fluctuator, which is the simultaneous flip of both spins in the same directions, for example going from $\ket{0,-1}$ to $\ket{+1,0}$. The processes $\ket{0, \pm 1}\bra{\mp 1,0}$ are resonant for $B=0$ and not anymore when the degeneracy between the $\ket{\pm 1}$ states is lifted by the magnetic field.
%
%We have seen however that for weak magnetic fields the eigenstates of each NV Hamiltonian are $\ket{\pm}$ and not $\ket{\pm 1}$, due to the transverse electric field. And the $\ket{+}$ and $\ket{-}$ states are not resonant in zero magnetic field. But we can see from Fig. \ref{simus ESR}-d) or \ref{ESR for T2*}-c) that the splitting $2 d_\perp E_\perp \approx 8\ \rm MHz$ is of the same order than the $1/T_1^{\rm dd}$ profile of linewidth $\Gamma^{\rm dd} \approx 8.8\ \rm MHz$ measured on Fig. [REF]. In zero external magnetic field, the $\ket{+}$ and $\ket{-}$ states are therefore close enough in energy that double flip can take place.
%
%To compute the depolarization induced by the double flips, we need to add another decay channel in the fluctuator model, on top of the flip-flop one. We also need to take into account the fact that the $\ket{+}$ and $\ket{-}$ states are not fully resonant, which will modify the $\bar{R}$ factor defined in [REF]. Finally, the question about the correlation  length of the electric field also needs to be asked.
%
%The results are summed up in Table \ref{table double flips}. We again give the results for the two extreme cases regarding the correlations in the electric field, and we show three possible scenario regarding the pseudo-resonance of the $\ket{+}$ and $\ket{-}$ states, represented by the $\bar{R}^2$ factor: $\bar{R}^2=1$ means that the states are fully resonant ($\omega_+-\omega_- \ll \Gamma^{\rm dd} $), this would be the maximum possible decay rate involving the double flips. $\bar{R}^2=0.5$ represents the case where the splitting between the states is equal to the fluctuator linewidth ($\omega_+-\omega_- \approx \Gamma^{\rm dd}$). This is close to the the experimental values we have. Finally $\bar{R}^2=0$ correspond to the case where the states are too far detuned for the double flips ($\omega_+-\omega_- \gg \Gamma^{\rm dd} $). We come back to the results of the last section where only considered the flip-flops.

\subsection{$T_2^*$ increase}

The last aspect we considered was the change in the ODMR lineshape discussed in sec. \ref{sec modif T2*} for weak magnetic field (we refer to the spectral linewidth as $T_2^*$ for conciseness).

To quantify the impact of the $T_2^*$ increase in the depolarization rate, we employ eq. [REF], where we use the values $2\gamma_f = 6.5\ \rm MHz$ estimated in the last chapter and we take for $\Gamma_f=\Gamma_{\rm NV}$ the values of half width at half maximum found in Fig. \ref{ESR for T2*}. We then find that the increase of $T_2^*$ in zero magnetic field occasions an increase of the decay rate by $\sim 9 \%$ for sample Sumi-2, and by $\sim 24 \%$ for sample CVD-pink.

The value shown in Fig. \ref{summary_theory} corresponds to a $9 \%$ increase.

In conclusion, we predict that the 4-classes degeneracy is by far the dominant causes for low field depolarization ($\sim \Gamma_1 \times 40$), followed by the double flips ($\sim \Gamma_1 \times 2$). We will however see in sec \ref{sec LFDM} that the double flips, along with the change of eigenbasis and increase in $T_2^*$ are determinant factor in the sensitivity of the low field magnetometry protocol.

In the following section we will confront the theoretical results with experimental ones.

\section{Experimental investigation of the low field depolarization}
\label{sec 4.3}
In this section we will try to experimentally discriminate the contribution of the four depolarization sources described in Fig. \ref{summary_theory} and confront the experimental values with the theoretical ones.

We will first determine the contribution of the 4-classes degeneracy by comparing [100] aligned and non-[100] aligned magnetic field in the weak magnetic field regime, before studying the influence of the double-flips thanks to purely transverse magnetic fields.

On top of the four previously discussed relaxation mechanisms, we have also considered two additional causes for the observation of low-field depolarization: the misalignment of the magnetic field (when $\mathbf{B}\parallel [100]$) and the influence of the laser polarization. Following further experiments detailed in appendix [REF], we have determined that these two effects were negligible compared to the four ones we identified before.

\subsection{NV-NV CR in the low magnetic field regime}

In this part, we will try to isolate the contribution of the 4-classes degeneracy from the three other ones by using the fact that all four classes can be at resonance for two distinct scenario:
\begin{itemize}
\item When $\mathbf{B}=0$.
\item When $\mathbf{B}\parallel [100]$ (cf. Fig. [REF] (la figure avec toutes les config)).
\end{itemize}

\subsubsection{Magnetic scan in a random direction}

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{Figures/scan_1x1x1x1}
\caption{Low field depolarization on sample ADM-150-1 for $\mathbf{B}$ in an arbitrary direction. a) ODMR spectrum for B=0. b) ODMR spectrum for B=60 G. c) PL contrast as a function of the external magnetic field. d) $1/T_1^{\rm dd}$ as a function of the external magnetic field. A $T_1$ measurement was recorded for each magnetic field value and fitted accorded to the protocol described in [REF] with $T_1^{\rm ph}=5\ \rm ms$}
\label{scan 1x1x1x1}
\end{figure}

First we will run a control experiment by scanning the magnetic field in a non [100] direction. Fig. \ref{scan 1x1x1x1}-a) and b) shows ODMR spectrum at the starting point ($|\mathbf{B}|=0$) and end point ($|\mathbf{B}|\approx 60\ \rm G$) of the magnetic field scan. The magnetic field orientation was chosen in order to maximize the splitting of the four classes.

Fig. \ref{scan 1x1x1x1}-c) and d) show the evolution of the NV PL and stretched lifetime $T_1^{\rm dd}$, defined in the last chapter [REF], as a function of the magnetic field. We can observe a decrease of $T_1^{\rm dd}$ by a factor of $\sim 10$ in zero magnetic field compared to non-zero magnetic field, with a half-width at half maximum of $\sim 10\ \rm G$. Correspondingly, we can observe a drop in PL by $\sim 4\%$ in zero magnetic field \footnote{We can also note the decrease in PL for $|\mathbf{B}|> 30\ \rm G$ is not corroborated by an increase of $1/T_1^{\rm dd}$. This is because this drop of PL is associated with transverse magnetic field state mixing, and not with spin depolarization.}.

We attribute the spin depolarization observed in Fig. \ref{scan 1x1x1x1}-d) to all four mechanisms described in the last part acting simultaneously. To determine the contribution of each mechanism individually, we will need to run dedicated experiments.

\subsubsection{Magnetic scan with $\mathbf{B} \parallel [100]$}

@@

This second experiment is similar to the last one except that the magnetic field was scanned in the [100] direction. Fig. \ref{scan 100}-a) and b) again show ODMR at the starting and ending point of the magnetic scan, while Fig. \ref{scan 100}-c) and d) show the PL and $1/T_1^{\rm dd}$ recorded during the scan.

The main difference with the previous experiment is that in this case, the degeneracy between the four classes is never lifted. We therefore attribute the increase in $1/T_1^{\rm dd}$ at low magnetic field to the three other identified mechanisms. The increase in the depolarization rate in zero field ($\sim 20\ \%$) as well as the decrease inPL ($\sim 0.5 \%$) are still present, but significantly smaller than in the previous case. We can note however that the half-width at half maximum of the zero field depolarization ($\sim 2\ \rm G$) is also smaller in this case.

We can also note a drop in PL and a corresponding increase of $1/T_1^{\rm dd}$ for $B \sim 20\ \rm G$ which corresponds to the NV$-^{13}$C-NV cross-relaxation discussed in [REF]. We can not explain however why the depolarization rate appear to be smaller at $B \sim 10\ \rm G$ than it is for $B > 30\ \rm G$.

The fact that the increase of the depolarization rate at zero field is about $\sim 50$ times weaker in Fig. \ref{scan 100} than it was in Fig. \ref{scan 1x1x1x1} confirms that the leading cause of depolarization in zero field is the degeneracy between the four classes of NV centers.

\subsection{NV-NV CR under purely transverse magnetic field}

Thank to the two previous experiments, we were able to discriminate the contribution of the 4-classes degeneracy from the three other mechanisms that are which are the eigenbasis change, the double-flips and the $T_2^* increase$. We now seek to discriminate the contribution of each of these three mechanisms.
